---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


```{r, cache=TRUE}
df <- read.csv("activity.csv")
df$dateString <- df$date
df$date <- as.POSIXct(df$date,format="%Y-%m-%d")
dfClean <- df[complete.cases(df),]
```

## What is mean total number of steps taken per day?

Steps per day
```{r countTable, results='asis', message=FALSE}
library(sqldf)
library(xtable)
stepsPerDay <- sqldf("select dateString as day,sum(steps) total_steps from dfClean group by date")   
xt <- xtable(stepsPerDay)
print(xt, type="html")

```

Steps per day histogram
```{r, message=FALSE}
library(ggplot2)
qplot(total_steps, data = stepsPerDay, main = 'Steps per Day')
```

Meand and median
```{r}
mean(stepsPerDay$total_steps)
median(stepsPerDay$total_steps)
```

## What is the average daily activity pattern?

```{r}
stepsPerInterval <- sqldf("select interval, avg(steps) mean_steps from dfClean group by interval")   

ggplot(data = stepsPerInterval, aes(interval, mean_steps )) + geom_line() + ylab("Average Number of Steps Per 5-Minute Interval") + xlab("Time Interval")

```

5-minute interval with the maximum average number of steps.
```{r}
idx <- which.max( stepsPerInterval$mean_steps )
stepsPerInterval$interval[[idx]]
```


## Imputing missing values
Missing values
```{r}
dfNAs <- df[!complete.cases(df),]
nrow(dfNAs)
```

Replacing missing values by the mean for that 5-minute intervals.

```{r}
for (i in 1:length(df[, 1])){
    if (is.na(df[i,1])){
        date <- df[i,2]
        interval <- df[i,3]
        mean_steps <- subset(stepsPerInterval, interval == df[i,3])$mean_steps
        df[i,1] <- mean_steps
    }

}

```


Steps per day histogram
```{r, message=FALSE}
stepsPerDay <- sqldf("select date as day,sum(steps) total_steps from df group by date") 
library(ggplot2)
qplot(total_steps, data = stepsPerDay, main = 'Steps per Day')
```

Meand and median
```{r}
mean(stepsPerDay$total_steps)
median(stepsPerDay$total_steps)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekenddays <- c("Saturday","Sunday")
areWeekends <- weekdays(df$date) %in% weekenddays

df$dayOfTheWeek <- rep(NA, nrow(df))
df[areWeekends,]$dayOfTheWeek <- "weekend"
df[!areWeekends,]$dayOfTheWeek <- "weekday"
df$dayOfTheWeek <- as.factor(df$dayOfTheWeek)


stepsPerIntervalPerDay <- sqldf("select interval, avg(steps) mean_steps, dayOfTheWeek from df group by interval,dayOfTheWeek")  

sp <- ggplot(data = stepsPerIntervalPerDay, aes(interval, mean_steps)) + geom_line() + ylab("Average Number of Steps Per 5-Minute Interval") + xlab("Time Interval")

sp + facet_grid(dayOfTheWeek ~ .)


```

